# Title
Operating Model for approximating the base case application of the Grym framework to the Antarctic krill fishery

# Subtitle
Description of input parameter values under scenario scn-1

# Author(s)
""


# Date
""

# Introduction

This is a supporting document for the application of the openMSE framework to the Antarctic Krill fishery. The primary objective of the analysis was to configure the openMSE components in a manner that closely approximates the methodology applied by the Grym framework [@maschette_etal20] for determining sustainable harvest rates for Krill. Specifically, the aim was to replicate the base-case implementation of Grym for the Krill stock as specified in [CCAMLR's code repository](https://github.com/ccamlr/Grym_Base_Case/tree/Simulations).

Here we focus on the specification of the main component of the openMSE framework, the Operational Model (OM), which sets up the parameters required to simulate the stock dynamics. In particular, we provide a description of the parameter values that underpin the OM for one of the input scenarios considered in the main analysis, referred to as scn-1. 

Further discussions and considerations regarding the configuration of the OMs to approximate the Grym framework are presented in the main document of the analysis. 


# Custom Parameters

### M and Perr

Simulation values for annual natural mortality (**M**) and recruitment process error (**Perr**) were obtained from a Proportional Recruitment (PR) model [@pavez_etal21] fitted to survey data specified under input scenario scn-1 (see main documentation for details on input parameters for the PR model). The PR method produces random draws of natural mortality $M$ and recruitment natural variability (expressed as CV, $CV_R$) expressing the uncertainty about those parameters. 

Generated draws of $M$ were used directly as simulation values of the OM parameter **M**. 

Draws of $CV_R$ were converted to log-normal standard deviations and plugged in as simulation values of **Perr** (i.e. $\texttt{Perr} = \sqrt{log(CV_R^2 + 1)}$).



### Age-plus group (plusgroup)

openMSE defaults to use the maximum age-class **maxage** as a plus group. However, in the Grym's base-case configuration, the final age class is not treated as an age-plus group. Therefore, to align with the Grym's specifications, the plus group option was switched off by setting the custom parameter as `OM@cpars$plusgroup <- 0`.



### Depletion optimization

The depletion optimization step in simulations for the historical period was switched off (`OM@cpars$qs <- rep(1, n_iter)`). This forces the derivation of the current depletion state to be based on selectivity and annual fishing mortality pattern. Therefore, values defined below for parameter **D** are ignored during simulations.


# Stock Parameters

<!-- ## Common_Name -->
<!-- No justification provided.  -->

## maxage

Value from @constable_delamare96.

In addition, as explained in section [Custom Parameters](#age-plus-group) above,
age-class 7 **is not** an age-plus group.

Thus, the assumption is that all individuals do not survive beyond the age of 7
years [@maschette_wotherspoon21].


## R0
Based on the value assumed for mean recruitment in Grym's base-case application.


## M
Simulation values for **M** provided in cpars (see [Custom
Parameters](#custom-parameters)). The text above on values being drawn from an
uniform distribution should be ignored.


## Msd
Assuming natural mortality in krill is constant over time, in alignment with
Grym's base-case application.


## h
Selected range of **h** values chosen to provide an approximation to Grym's
approach of simulating annual recruitment. In Grym's base-case, simulated number
of recruits are *mostly* independent of stock size, unless when they are
penalised by a depletion factor if the stock size in the preceding year falls
below 20% of its pre-exploitation levels.

Specifying high values of parameter **h**  results in strongly concave
stock-recruitment curves, characterized by a steep increase in recruitment at
low levels of stock size that approaches an asymptote. This implies that the
simulated values of recruitment will be nearly independent of stock size, except
when the stock is at 20% of its pre-exploitation size or below.


## SRrel
Grym's base-case does not incorporate density dependence in recruitment success,
making the Ricker model less suitable as an approximation. Instead, the
Beverton–Holt stock-recruitment model, with the values for parameter **h**
specified as above, is expected to provide a better approximation to the Grym
approach.


## Perr
Simulation values of **Perr** provided in cpars (see [Custom
Parameters](#custom-parameters)). The text above about values being drawn from
an uniform distribution should be ignored.


## AC
Assuming no autocorrelation between consecutive annual recruitments.



## Linf
Value from @constable_delamare96, in millimeters. As in Grym's base case,
the $L_\infty$ parameter is treated as known (i.e. no uncertainty).


## Linfsd
Assuming no inter-annual variation in $L_\infty$.


## K
Value from @thanassekos_etal21. Parameter $K$ is assumed to be known without
error.


## Ksd
Assuming no inter-annual variation in growth parameter $K$. 

## t0
Value from @constable_delamare96. As in Grym's base-case, age at size-0 assumed
as known (i.e. no uncertainty).


## LenCV

Grym's base case does not account for variation in length-at-age, implying that
all individuals follow exactly the average growth curve. However, trial runs in
openMSE indicated that simulations would crash if variation in length-at-age was
set to zero.

Hence, to avoid simulation crashes, a very low level of variation (CV of 1%) in
length-at-age was assumed in openMSE simulations.

Furthermore, variation in length-at-age treated as known without error, with
fixed values across simulations.


## L50
Same values as minimum and maximum length (mm) at 50% maturity as specified
under scenario scn-1, originally provided in @maschette_etal21 (check main
analysis document for further details).

## L50_95

Grym's base-case formulates size-at-maturity using the ogive ramp function
specified in terms `L50` and the ramp's width `matrange` (i.e. the range at
which maturity occurs).

Thus, `L95` (length at which 95% of individuals are matures) can be derived via
the quantile function of the uniform distribution with limits `L50 +- matrange/2`.

For a (known) width of 6mm, as specified under input scenario scn-1, we derive the `L50_95`
parameter value as follows:

```{r}
matrange <- 6
L50 <- 35 # arbitrary as `matrange` is treated as independent of midpoint L50 in Maschette et al (2021)
L95 <- qunif(0.95, L50 - matrange/2, L50 + matrange/2)
L95 - L50 
```
Similarly to the Grym approach, where `matrange` is treated as known, **L50_95**
is also kept fixed across simulations.


## D
Arbitrary values as depletion optimization step is not carried out (see
[Depletion optimization](#depletion-optimization) section above).


## a
Value from @sc-camlr00. As in Grym's base-case, the parameter **a** is assumed
to be known without error (i.e. fixed values across simulations).

## b
Value from @sc-camlr00. Parameter **b** assumed to be known without error (i.e.
fixed values across simulations).


## Size_area_1
Parameter value not allowed to be 1 (i.e. forcing one single area), so
specifying that the two sub-areas have equal size.

## Frac_area_1
Parameter value not allowed to be 1 (i.e. forcing whole population to a single
area), so specifying that population is evenly split between the two sub-areas.


## Prob_staying
Assuming 50% chance of movement between sub-areas.

## Fdisc
Zero discard mortality rate, as discards are assumed non-existent in the Krill
fishery, similarly to the Grym's base-case application.


## Source
""


# Fleet Parameters

## nyears
Minimum number of years allowed for the historical period in the current version
of openMSE/MSEtools. Any value lower than 2 will cause simulation errors.


## Spat_targ
Assuming fishing effort is allocated across areas in proportion to the
population density in each area.


## EffYears
Arbitrary value. At least a single value required, otherwise the simulation will
crash.


## EffLower
Stock assumed to remain unexploited during the historical period.


## EffUpper
Stock assumed to remain unexploited during the historical period.

## Esd
Population assumed to remain unexploited during the historical period, so
variability in historical fishing mortality is not applicable.

## qinc
Assuming no change in mean gear efficiency over the projection period.


## qcv
Assuming no inter-annual variability in gear efficiency over projection years.


## L5

Analogously to the maturity ogive, the Grym approach specifies gear selectivity
as a ramp function governed by two parameters: (i) the ramp's midpoint `L50`,
expressing the length at which 50% of the fish are vulnerable to selection by
the fishery; and (ii) `selrange`, the width of the ramp.

Minimum and maximum values of `L50` and `selrange` are provided in
@thanassekos_etal21, which are used to derive the minimum and maximum values of
`L5` via the uniform quantile function, as follows:

```{r}
selrange <- 11
L50 <- c(30, 35)
L5 <- qunif(0.05, L50 - selrange/2, L50 + selrange/2)
L5
```


## LFS

Values derived from `L50` and `selrange`, as above.

```{r}
L50 + selrange/2
```


## Vmaxlen
All individuals considered to be selected at $L_{\infty}$.


## isRel
Selectivity parameters are in absolute units.

## LR5
Analogously to the Grym's base case, assuming no discards - i.e. all individuals
selected by the gear are retained.



## LFR
Idem to **LR5** - Assuming no discards and hence all individuals selected by the
gear are retained.


## Rmaxlen
Same as **Vmaxlen**, i.e. 100% of fish are retained at $L_{\infty}$.

## DR
Assuming no discards.


## CurrentYr
Arbitrary value in the context of the openMSE-Grym approximation. Choosing the
starting year of the project.


## MPA
Both areas assumed to be open to fishing at all times.



# Obs Parameters

This section is not applicable in the context of the openMSE-Grym approximation, as the Management Procedures (MPs) that will be considered for the analysis (i.e. one MP for each of the considered $\gamma$ values) don't require sampled data from the fishery. Therefore, all Obs parameters will be set to zero.


## Cobs
Catch assumed known without observation error.

## Cbiascv
Catch assumed known without observation bias.

## CAA_nsamp
No catch-at-age sampling.

## CAA_ESS
No catch-at-age sampling.

## CAL_nsamp
No catch-at-length sampling.

## CAL_ESS
No catch-at-length sampling.

## Iobs
No observation error in the relative abundance index.

## Btobs
No observation error in the absolute abundance index.

## Btbiascv
No bias in observations of stock biomass.

## beta
Assuming no hyperstability/hyperdepletion in abundance estimation.

## LenMbiascv
Assuming no bias in estimated biological parameters.

## Mbiascv
Assuming no bias in estimated biological parameters.

## Kbiascv
Assuming no bias in estimated biological parameters.

## t0biascv
Assuming no bias in estimated biological parameters.

## Linfbiascv
Assuming no bias in estimated biological parameters.

## LFCbiascv
Assuming no bias in estimated biological parameters.

## LFSbiascv
Assuming no bias in estimated biological parameters.

## FMSY_Mbiascv
Assuming no bias.

## BMSY_B0biascv
Assuming no bias.

## Irefbiascv
Assuming no bias.

## Brefbiascv
Assuming no bias.

## Crefbiascv
Assuming no bias.

## Dbiascv
Assuming no bias.

## Dobs
Assuming no bias.

## hbiascv
Assuming no bias.

## Recbiascv
Assuming no bias.

## sigmaRbiascv
Assuming no bias.

## Eobs
No observation error around the total effort

## Ebiascv
Assuming no bias.



# Imp Parameters

Similarly to the Grym's base-case approach, it is assumed that the recommended
Total Allowable Catches (TACs) under considered MPs are enforced perfectly,
meaning that catches never exceed nor fall short of the stipulated quotas.

## TACFrac
Assuming perfect implementation of TAC recommendations.

## TACSD
Assuming perfect implementation of TAC recommendations.

## TAEFrac
Assuming perfect implementation - not applicable for the openMSE-Grym approximation.

## TAESD
Assuming perfect implementation - not applicable for the openMSE-Grym approximation.

## SizeLimFrac
Assuming perfect implementation of size limit rules - not applicable for the
openMSE-Grym approximation.

## SizeLimSD
Assuming perfect implementation of size limit rules - not applicable for the
openMSE-Grym approximation.


# References

